### 2025-03-11 paper/preprint/paper.tex
Add two points in 5.2 limitations

\usepackage{arxiv}
\usepackage{orcidlink}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\usepackage{placeins}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\usepackage{amsthm}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
\theoremstyle{plain}
\newtheorem{corollary}{Research Question}[section]
\theoremstyle{plain}
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{remark}
\AtBeginDocument{\renewcommand*{\proofname}{Proof}}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\newtheorem{refremark}{Remark}[section]
\newtheorem{refsolution}{Solution}[section]
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

@@

\textbf{\emph{CT increases the training time of models.}}

\textbf{\emph{Immutable features may have proxies.}}

\textbf{\emph{Interventions on features may have downstream impacts on
fairness.}}

@@

>>>>

\usepackage{arxiv}
\usepackage{orcidlink}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\usepackage{placeins}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\usepackage{amsthm}
\theoremstyle{plain}
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
\theoremstyle{plain}
\newtheorem{corollary}{Research Question}[section]
\theoremstyle{remark}
\AtBeginDocument{\renewcommand*{\proofname}{Proof}}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\newtheorem{refremark}{Remark}[section]
\newtheorem{refsolution}{Solution}[section]
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

@@

\textbf{\emph{CT increases the training time of models.}} Counterfactual
training promotes explainability through CEs and robustness through AEs
at the cost of longer training times compared to conventional training
regimes. While higher numbers of iterations and counterfactuals per
iteration positively impact the quality of found solutions, they also
increase the required amount of computations. We find that relatively
small grids with 270 settings can take almost four hours for more
demanding datasets on a high-performance computing cluster with 34 2GB
CPUs (see details in \textbf{?@sec-hardware}). However, there are three
factors that attenuate the impact of this limitation. First, CT provides
counterfactual explanations for the training samples essentially for
free, which may be beneficial in many ADM systems. Second, we find that
CT can retain its value when used as a ``fine-tuning'' training regime
for conventionally-trained models. Third, in principle, CT yields itself
to multi-threaded execution.

\textbf{\emph{Immutable features may have proxies.}} In
Proposition~\ref{prp-mtblty} we define an approach to protect immutable
features and thus increase the actionability of the generated
counterfactuals. However, this approach requires that model owners
define the mutability constraints for (all) features considered by the
model. Even with sufficient domain knowledge to protect all immutable
features---ones that cannot be changed at all and ones that cannot be
reasonably expected to change---there may exist proxies that are
theoretically mutable (and hence should not be protected) but preserve
enough information about the principals to counteract the protections.
As one example, consider the Adult dataset used in our experiments where
the mutable education status is a proxy for the immutable age, in that
the attainment of degrees is correlated with age. Delineating
actionability is a major undecided challenge in the AR literature (see,
e.g.,
\citeproc{ref-venkatasubramanian2020philosophical}{Venkatasubramanian
and Alfano 2020}) impacting the capacity of CT to increase the
explainability of the model.

\textbf{\emph{Interventions on features may have downstream impacts on
fairness.}}\\
- \href{https://arxiv.org/abs/2401.16088}{Equality of opportunity}?\\
- \href{https://proceedings.mlr.press/v202/gao23d/gao23d.pdf}{Social
segregation}?\\
- Supporting AR within one context/system may still unfairly target
recourse-affected individuals within other contexts/systems? In Example
3.1 younger individuals gain access to a loan but their age could still
lead to, e.g., enforcing stronger supervisory mechanisms?

@@

\bibitem[\citeproctext]{ref-venkatasubramanian2020philosophical}
Venkatasubramanian, Suresh, and Mark Alfano. 2020. {``The Philosophical
Basis of Algorithmic Recourse.''} In \emph{Proceedings of the 2020
Conference on Fairness, Accountability, and Transparency}, 284--93. FAT*
'20. New York, NY, USA: Association for Computing Machinery.
\url{https://doi.org/10.1145/3351095.3372876}.

### 2025-03-11 paper/sections/discussion.qmd
Add two points in 5.2 limitations

***CT increases the training time of models.***

***Immutable features may have proxies.***

***Interventions on features may have downstream impacts on fairness.***

>>>>

***CT increases the training time of models.*** Counterfactual training promotes explainability through CEs and robustness through AEs at the cost of longer training times compared to conventional training regimes. While higher numbers of iterations and counterfactuals per iteration positively impact the quality of found solutions, they also increase the required amount of computations. We find that relatively small grids with 270 settings can take almost four hours for more demanding datasets on a high-performance computing cluster with 34 2GB CPUs (see details in @sec-hardware). However, there are three factors that attenuate the impact of this limitation. First, CT provides counterfactual explanations for the training samples essentially for free, which may be beneficial in many ADM systems. Second, we find that CT can retain its value when used as a "fine-tuning" training regime for conventionally-trained models. Third, in principle, CT yields itself to multi-threaded execution.

***Immutable features may have proxies.*** In @prp-mtblty we define an approach to protect immutable features and thus increase the actionability of the generated counterfactuals. However, this approach requires that model owners define the mutability constraints for (all) features considered by the model. Even with sufficient domain knowledge to protect all immutable features---ones that cannot be changed at all and ones that cannot be reasonably expected to change---there may exist proxies that are theoretically mutable (and hence should not be protected) but preserve enough information about the principals to counteract the protections. As one example, consider the Adult dataset used in our experiments where the mutable education status is a proxy for the immutable age, in that the attainment of degrees is correlated with age. Delineating actionability is a major undecided challenge in the AR literature [see, e.g., @venkatasubramanian2020philosophical] impacting the capacity of CT to increase the explainability of the model.

***Interventions on features may have downstream impacts on fairness.***  
- [Equality of opportunity](https://arxiv.org/abs/2401.16088)?  
- [Social segregation](https://proceedings.mlr.press/v202/gao23d/gao23d.pdf)?  
- Supporting AR within one context/system may still unfairly target recourse-affected individuals within other contexts/systems? In Example 3.1 younger individuals gain access to a loan but their age could still lead to, e.g., enforcing stronger supervisory mechanisms?
